---
title: "Simulation study for debiased learning"
header-includes:
  - \usepackage{statstyle}
output:
  pdf_document: 
    keep_tex: true
---

```{r setup, include=FALSE}
N = 2000
n = 40
m = 20
p = 15
s = 10

set.seed(2)

X = matrix(rnorm(N * p, 2, 1), nr = N, nc = p)
beta = c(rep(1, s), rep(0, p - s))
e = rnorm(N, 0, 0.1)
y = X %*% beta + e 
t_y = sum(y)

# pi1 = e^2 / sum(e^2) * n

r = 0.75

ratio = (1 - r) / r
z = rnorm(n = N, 0, sqrt(ratio)) + e
z_sorted_idx = order(z)
len = round(N / 4)

# pi1 = rep(n / N, N)
# pi1 = rep(n / N, N) + rnorm(N, 0, 0.0025)
# pi1 = rep(n / N, N) + e * n / N * 0.2

# pi1 = (y - min(y)) / (max(y) - min(y)) / 2 + 0.25
# pi1 = (e - min(e)) / (max(e) - min(e)) / 2 + 0.10
# pi1 = pi1 / sum(pi1) * n
# d1 = 1 / pi1
```

# Simulation setup

Let $A$ be a sample of size $n =$ `r n` selected from the population $U$ of size $N =$ `r N`. Let $y$ be a variable of interest and we try to estimate the population total $t_y = \sum_{i \in U} y_i$ and $\bf x_{i}$ of size $p$ be the auxiliary variables associated with $y$ where $x_{ij} \iid N(2, 1)$. Suppose the following superpopulation model holds
$$
y_i = \bf x_i^T \bf \beta + \varepsilon_i, \quad \varepsilon_i \iid N(0,\sigma^2), i = 1, \cdots, N
$$
where $\sigma^2 = 1$ and $\bf \beta = (1, \cdots, 1, 0, \cdots, 0)$ with $s =$ `r s` 1s and $p - s =$ `r p - s` 0s. We assume the samples are drawn from the SRS sampling.

We consider the following estimators.
\begin{itemize}
  \item Difference estimator
  $$
  \hat t_y = \pa{\sum_{i \in U}\bf x_i}^T \bf \beta + \sum_{i \in A} d_i \pa{y_i - \bf x_i^T \bf \beta}
  $$
  \item GREG estimator
  $$
  \hat t_y = \pa{\sum_{i \in U}\bf x_i}^T \hat{\bf \beta} + \sum_{i \in A} d_i \pa{y_i - \bf x_i^T \hat{\bf \beta}}
  $$
  where $\hat {\bf \beta} = \pa{\sum_{i \in A} d_i \bf x_i \bf x_i^T}^{-1} \pa{\sum_{i \in A} d_i \bf x_i y_i}$
  
  \item Model-based estimator
  $$
  \hat t_y = \pa{\sum_{i \in U}\bf x_i}^T \hat{\bf \beta}
  $$
  
  \item Debiased GREG estimator(Proposed)
    $$
  \hat t_y = \pa{\sum_{i \in U}\bf x_i}^T \pa{ \frac{\sum_{i \in A_2}d_i}{\sum_{i \in A}d_i} \hat{\bf \beta}_1 + \frac{\sum_{i \in A_1}d_i}{\sum_{i \in A}d_i} \hat{\bf \beta}_2} + \sum_{i \in A_1} d_i \pa{y_i - \bf x_i^T \hat{\bf \beta}_2} + \sum_{i \in A_2} d_i \pa{y_i - \bf x_i^T \hat{\bf \beta}_1}
  $$
  where $A_1$ and $A_2$ are the subsamples of $A$ with $A_1 \bigcup A_2 = A$, and $\hat {\bf \beta}_1$ and $\hat {\bf \beta}_2$ are the regression coefficients built upon the sample $A_1$ and sample $A_2$ respectively.
  \item Design unbiased GREG estimator(Sande, 2021)
  $$
  \hat t_y = \pa{\sum_{i \in U}\bf x_i}^T \hat{\bf \beta}_1 + \sum_{i \in A_1} \pa{y_i - \bf x_i^T \hat{\bf \beta}_1} + \sum_{i \in A_2} \frac{1}{P(i \in A_2 | i \in A)} \pa{y_i - \bf x_i^T \hat{\bf \beta}_1}
  $$
  
  %\item LASSO estimator
  %\item Debiased Lasso estimator(Zhang \& Zhang, 2014)
\end{itemize}

```{r include=FALSE}
res = NULL
res2 = NULL
res3 = NULL
SIMNUM = 500
X = cbind(1, X)
beta = c(0, beta)
for(simnum in 1:SIMNUM){
  print(simnum)
  # set.seed(simnum)
  
  Index = rep(0, n)
d1 = rep(0, N)
n_h = c(15, 20, 30, 35) / 100 * n
cumn_h = cumsum(n_h)
for (i in 1:4){
  Idx_z = (len * (i - 1)+ 1): (len * i)
  Idx = z_sorted_idx[Idx_z]
  d1[Idx] = len / n_h[i]
  from = ifelse(i == 1, 0, cumn_h[i-1])
  Index[(from + 1) : cumn_h[i]] = sample(Idx, size = n_h[i], replace = FALSE)
}

  
  # Index = sample(1:N, size = n, replace = FALSE, prob = pi1)
y_s = y[Index]
X_s = X[Index, ,drop = F]
d1_s = d1[Index]
lm_obj = lm(y_s ~  0 + X_s, weights = d1_s)
beta_hat = lm_obj$coefficients
#drop(solve(t(X_s) %*% diag(d1_s) %*% X_s, t(X_s) %*% diag(d1_s) %*% y_s))

# beta_hat = unname(lm(y_s ~  t(apply(X_s, 1, function(k) k 
# - colSums(X_s * d1_s) / N)), weights = d1_s)$coefficients[-1])

y_HT = sum(y_s * d1_s)

y_diff = drop(colSums(X) %*% beta + sum((y_s - drop(X_s %*% beta)) * d1_s))

y_GREG = drop(colSums(X) %*% beta_hat + sum((y_s - drop(X_s %*% beta_hat)) * d1_s))

y_model = drop(colSums(X) %*% beta_hat)

Index_sub1 = Index[sample(1:n, size = m, replace = FALSE)]
Index_sub2 = Index[!(Index %in% Index_sub1)]

y_s1 = y[Index_sub1]
X_s1 = X[Index_sub1, ,drop = F]
d1_s1 = d1[Index_sub1]
lm_obj1 = lm(y_s1 ~  0 +X_s1, weights = d1_s1)
beta_hat1 = lm_obj1$coefficients

y_s2 = y[Index_sub2]
X_s2 = X[Index_sub2, ,drop = F]
d1_s2 = d1[Index_sub2]
lm_obj2 = lm(y_s2 ~ 0 +  X_s2, weights = d1_s2)
beta_hat2 = lm_obj2$coefficients

# y_debiased = drop(colSums(X) %*% beta_hat + sum((y_s1 - drop(X_s1 %*% beta_hat2)) * d1_s1) 
# + sum((y_s2 - drop(X_s2 %*% beta_hat1)) * d1_s2))

y_debiased = drop(colSums(X) %*% (sum(d1_s2) / sum(d1_s) * beta_hat1 + sum(d1_s1) / sum(d1_s) * beta_hat2) + 
                    sum((y_s1 - drop(X_s1 %*% beta_hat2)) * d1_s1) + sum((y_s2 - drop(X_s2 %*% beta_hat1)) * d1_s2))

y_unbiased = drop(colSums(X) %*% beta_hat1 + sum((y_s1 - drop(X_s1 %*% beta_hat1))) 
                  + sum((y_s2 - drop(X_s2 %*% beta_hat1)) * 2))

Omega = matrix(1 - n / (n-1) * (N-1) / N, nr = n, nc = n)
diag(Omega) = 1 - n / N


sigma_HT = sqrt(drop(t(y_s)  %*% Omega %*% y_s * N^2 / n^2))

e_s = y_s - X_s %*% beta
sigma_diff = sqrt(drop(t(e_s)  %*% Omega %*% e_s * N^2 / n^2))

# sum(sapply(1:n, function(k) sapply(1:n, function(l) 
#   ifelse(k == l, 1 - n / N, 1 - n / (n - 1) * (N - 1) / N) * e_s[k] * e_s[l] * N / n * N / n  ))) / N^2

e_s = y_s - X_s %*% beta_hat
sigma_GREG = sqrt(drop(t(e_s)  %*% Omega %*% e_s * N^2 / n^2))

# e_s = ifelse(Index %in% Index_sub1, y_s - X_s %*% beta_hat2, y_s - X_s %*% beta_hat1)
# # e_s = ifelse(Index %in% Index_sub1, y_s - X_s %*% beta_hat1, y_s - X_s %*% beta_hat2)
# # e_s = c(y_s1 - X_s1 %*% beta_hat2, y_s2 - X_s2 %*% beta_hat1)
# sigma_Debiased = sqrt(drop(t(e_s)  %*% Omega %*% e_s * N^2 / n^2))

e_s1 = ifelse(Index %in% Index_sub1, y_s - X_s %*% beta_hat1, 0)
e_s2 = ifelse(Index %in% Index_sub2, y_s - X_s %*% beta_hat2, 0)

sigma_Debiased = sqrt(drop((t(e_s1)  %*% Omega %*% e_s1 + t(e_s2)  %*% Omega %*% e_s2) * N^2 / n^2))

that_k = sapply(1:n, function(k)  sum((y_s[-k]) * d1_s[-k]) * (n) / (n-1))
sigma_HT2 = sqrt(sum((that_k- y_HT)^2) * (n - 1) / n)

that_k = sapply(1:n, function(k)  sum((y_s[-k]) * d1_s[-k]) * (n) / (n-1)  + drop((colSums(X) - colSums(X_s[-k,,drop = F] * d1_s[-k]) * (n) / (n-1)) %*% beta))

sigma_diff2 = sqrt(sum((that_k- y_diff)^2) * (n - 1) / n)

that_k = sapply(1:n, function(k)  sum((y_s[-k]) * d1_s[-k]) * (n) / (n-1)  + drop((colSums(X) - colSums(X_s[-k,,drop = F] * d1_s[-k]) * (n) / (n-1)) %*% drop(solve(t(X_s[-k,,drop = F]) %*% diag(d1_s[-k]) %*% X_s[-k,,drop = F], t(X_s[-k,,drop = F]) %*% diag(d1_s[-k]) %*% y_s[-k]))))

# that_k = sapply(1:n, function(k)  sum((y_s) * d1_s)  + drop((colSums(X) - colSums(X_s * d1_s)) %*% drop(solve(t(X_s[-k,]) %*% diag(d1_s[-k]) %*% X_s[-k,], t(X_s[-k,]) %*% diag(d1_s[-k]) %*% y_s[-k]))))
    
sigma_GREG2 = sqrt(sum((that_k- y_GREG)^2) * (n - 1) / n)

that_k1 = sapply(1:length(d1_s1), function(k) {betahat_k1 = drop(solve(t(X_s1[-k,,drop = F]) %*% diag(d1_s1[-k]) %*% X_s1[-k,,drop = F], t(X_s1[-k,,drop = F]) %*% diag(d1_s1[-k]) %*% y_s1[-k]))
drop(colSums(X) %*% (sum(d1_s2) / sum(d1_s) * betahat_k1 + sum(d1_s1) / sum(d1_s) * beta_hat2) + 
                    sum((y_s1[-k] - drop(X_s1[-k,,drop = F] %*% beta_hat2)) * d1_s1[-k]) + sum((y_s2 - drop(X_s2 %*% betahat_k1)) * d1_s2))
})

that_k2 = sapply(1:length(d1_s2), function(k) {betahat_k2 = drop(solve(t(X_s2[-k,,drop = F]) %*% diag(d1_s2[-k]) %*% X_s2[-k,,drop = F], t(X_s2[-k,,drop = F]) %*% diag(d1_s2[-k]) %*% y_s2[-k]))
drop(colSums(X) %*% (sum(d1_s2) / sum(d1_s) * beta_hat1 + sum(d1_s1) / sum(d1_s) * betahat_k2) + 
                    sum((y_s1 - drop(X_s1 %*% betahat_k2)) * d1_s1)) + sum((y_s2[-k] - drop(X_s2[-k,,drop = F] %*% beta_hat1)) * d1_s2[-k])
})

that_k_null = rep(0, length(that_k))
that_k_null[Index %in% Index_sub1] = that_k1
that_k_null[Index %in% Index_sub2] = that_k2

sigma_Debiased2 = sqrt(sum((that_k_null - y_debiased)^2) * (n - 1) / n)

res = rbind(res, c(Diff = y_diff, GREG = y_GREG, Model = y_model, Debiased = y_debiased, Unbiased = y_unbiased, HT = y_HT))

res2 = rbind(res2, c(HT = sigma_HT, Diff = sigma_diff, GREG = sigma_GREG, Debiased = sigma_Debiased, HT2 = sigma_HT2, Diff2 = sigma_diff2, GREG2 = sigma_GREG2, Debiased2 = sigma_Debiased2))

# res3 = rbind(res3, c(sum((y_s1 - drop(X_s1 %*% beta_hat2)) * d1_s1), sum((y_s2 - drop(X_s2 %*% beta_hat1)) * d1_s2)))

res3 = rbind(res3, c(colMeans(X_s), beta_hat))

}

BIAS = colMeans(res - t_y)
SE = apply(res, 2, function(x) sqrt(var(x) * (length(x)-1)/length(x) ))
RMSE = apply(res - t_y, 2, function(x) sqrt(mean(x^2)))
```

```{r}
cbind(BIAS, SE, RMSE)
```


# Variance estimation
Consider the following variance estimation method for estimating the variance of debiased estimator:
$$
V_p\pa{\sum_{i \in A_1}\frac{y_i - \bf x_i \hat {\bf \beta}_2}{\pi_i}} + V_p\pa{\sum_{i \in A_2}\frac{y_i - \bf x_i \hat {\bf \beta}_1}{\pi_i}}
$$

```{r}
BIAS2 = colMeans(res2 - rep(SE[rep(c(6,1,2,4),2)], each = SIMNUM))
SE2 = apply(res2, 2, function(x) sqrt(var(x) * (length(x)-1)/length(x) ))
RMSE2 = apply(res2 - rep(SE[rep(c(6,1,2,4),2)], each = SIMNUM), 2, function(x) sqrt(mean(x^2)))
MidErr2 = apply(res2 - rep(SE[rep(c(6,1,2,4),2)], each = SIMNUM), 2, median)

cbind(BIAS2, SE2, RMSE2, REl_BIAS = BIAS2 / SE[rep(c(6,1,2,4),2)])
cbind(BIAS2, REl_BIAS = BIAS2 / SE[rep(c(6,1,2,4),2)])
```

```{r}
plot(res2[,2], main = "GREG", ylim = c(min(res2[,2]), SE[2]))
abline(h = SE[2], col = "blue")
abline(h = median(res2[,2]), col = "red")
```



```{r}
plot(res2[,3], main = "Debiased")
abline(h = SE[4], col = "blue")
abline(h = median(res2[,3]), col = "red")
# median(res2[,3])
```

```{r}
# hist(res2[,3], main = "SE of Debiased estimator")
library(ggplot2)
ggplot() + 
  geom_density(data = data.frame(res), aes(x = Diff, fill = "Diff"), alpha = 0.3) +
  geom_density(data = data.frame(res), aes(x = GREG, fill = "GREG"), alpha = 0.3) + 
  geom_density(data = data.frame(res), aes(x = Debiased, fill = "Debiased"), alpha = 0.3) + 
  geom_vline(xintercept = t_y, col = "red")
```

